---
title: "706 Midterm 1_SAP"
author: "Linke Li"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(here)
library(readr)
library(table1)
library(ggplot2)
library(tidyverse)
library(mice)
library(table1)
library(flextable)
library(knitr)

# clean environment before each project
rm(list = ls())
```

# Read in and check Datasets

```{r}
HR_raw <- read.csv(here("frmgham2.csv"))
sum(is.na(HR_raw)) # That's lot of NA D:
table(HR_raw$TIME)
ggplot(HR_raw, aes(x = TIME)) +
  geom_bar()

# creating a clean dataset
HR_rawshort <- HR_raw %>%
  select(RANDID, SEX, AGE, DIABETES, HEARTRTE, TIME, PERIOD)
sum(is.na(HR_rawshort)) #only 6 missing variable in the short dataset

length(unique(HR_rawshort$RANDID)) #4434 patient in the dataset

# check the rows with missing values -- I don't think we should remove NA? mice imputation?
subset(HR_rawshort, is.na(HR_rawshort$HEARTRTE))

# mice imputation is used to generate 10 sets of heart rate data, 
# and the mean of that 10 sets are used to replace the heart rate.
set.seed(66)
HRMI <- mice(HR_rawshort, method = "pmm", m = 10, maxit = 20)
HRMI_final <- complete(HRMI)

length(unique(HRMI_final$RANDID))

#Check for abnormal data by uisng min and max
min(HRMI_final$HEARTRTE) #37
max(HRMI_final$HEARTRTE) #220

min(HRMI_final$SEX)
max(HRMI_final$SEX)

min(HRMI_final$DIABETES)
max(HRMI_final$DIABETES)

min(HRMI_final$TIME)
max(HRMI_final$TIME)

min(HRMI_final$AGE)
max(HRMI_final$AGE)

```
### table 1 - summary statistics
```{r table-1}
# Table 1 package
label(HRMI_final$RANDID) <- "Patient random ID"
label(HRMI_final$AGE) <- "Age"
HRMI_final$SEX <- factor(HRMI_final$SEX, levels=c(1, 2), labels=c("Men", "Women"))
HRMI_final$DIABETES <- factor(HRMI_final$DIABETES, levels=c(0, 1), labels=c("No Diabetes", "Diabetes"))
label(HRMI_final$HEARTRTE) <- "Heart Rate"
label(HRMI_final$TIME) <- "Time at Exam"
label(HRMI_final$PERIOD) <- "Exam Period"


units(HRMI_final$AGE) <- "Years"
units(HRMI_final$HEARTRTE) <- "beats per minute(bpm)"
units(HRMI_final$TIME) <- "Days"

#Specific render for display quantiles
render.median.Q <- function(x, ...) {
  c('', 
    `Mean (SD)` = sprintf("%s (%s)", round(mean(x), 2), round(sd(x), 2)),
    `Median [Q1,Q3]` = sprintf("%s [%s, %s]", median(x), 
                             quantile(x, 0.25), quantile(x, 0.75)))
}

#display table 1
Dat_Table1 <- table1(~ AGE + SEX + DIABETES + HEARTRTE | PERIOD, data = HRMI_final,overall = "Total")
Dat_Table1

# Convert to flextable because I have no idea why it doesn't knit into a word
 t1flex(Dat_Table1) %>% 
   save_as_docx(path="Table1_706.docx")
```


### Aim1:
describe the distribution of heart rate and diabetes and test for the association of heart rate and diabetes in Period 1 of the study.
```{r aim1}
# the association between heart rate and diabetes for patients during period 1
HRMI_a1 <- HRMI_final %>%
  filter(PERIOD == 1)

# distribution of Heart rate
ggplot(HRMI_a1,
       aes(x = HEARTRTE)) +
  geom_bar() +
  xlab("Heart Rate(bpm)") +
  ggtitle("Distribution of Heart Rate in the Study Population during Period 1")

summary(HRMI_a1$HEARTRTE)
sd(HRMI_a1$HEARTRTE)

# normality of heart rate assumption?
qqnorm(HRMI_a1$HEARTRTE, pch = 1)
qqline(HRMI_a1$HEARTRTE, col = "steelblue", lwd = 2)

# Distribution of Diabete status
ggplot(HRMI_a1,
       aes(x = DIABETES, color = DIABETES)) +
  geom_bar() +
  xlab("Diabete Status") +
  ggtitle("Distribution of Diabetes in the Study Population during Period 1")

sum(ifelse(HRMI_a1$DIABETES == 0, 1, 0))
sum(ifelse(HRMI_a1$DIABETES == 1, 1, 0))

#logistic rgeression for association
a1model <- glm(DIABETES ~ HEARTRTE,family=binomial(link='logit'),data = HRMI_a1)
summary(a1model)
a1model
```
p-value: 0.000863
The result logistic regression model: $\hat{y} = -5.375 + 0.023 Heart Rate$
### Aim 2
Aim 1 latex: The Diabetes outcome in period 1 will be a binary dummy variable with values either 0 or 1. Heart rate in period 1 will take on real-number values between 44 and 143.    

$E[Y_i] = \pi_i$  
$(\pi_i) = log(\frac{\pi_i}{1-\pi_i}) = \beta_0 + \beta_1 x_i^{Heart Rate}$

```{r aim 2}
# The model fitted
a2model <- glm(DIABETES ~ HEARTRTE + AGE + HEARTRTE : AGE, family=binomial(link='logit'),data = HRMI_a1)
summary(a2model)
a2model

a2model1 <- glm(DIABETES ~ HEARTRTE + AGE, family=binomial(link='logit'),data = HRMI_a1)
summary(a2model1)
a2model1

# confirm range of age groups
min(HRMI_a1$AGE)
max(HRMI_a1$AGE)
ggplot(HRMI_a1, aes(AGE)) +
  geom_bar()

# upon checking, the minimal age is 32 and the maximum is 70.
# The same time upon checking the bar graph without any categorization of the continuous variable age,
# 5 groups/bins seems to be a reasonable choice.


# Catgorize age into sub groups
HRMI_a2 <- HRMI_a1 %>%
  mutate(CatAge = NA)
HRMI_a2$CatAge[HRMI_a2$AGE < 40] <- "Below 40"
HRMI_a2$CatAge[40 <= HRMI_a2$AGE & HRMI_a2$AGE < 50] <- "40s"
HRMI_a2$CatAge[50 <= HRMI_a2$AGE & HRMI_a2$AGE < 60] <- "50s"
HRMI_a2$CatAge[60 <= HRMI_a2$AGE & HRMI_a2$AGE < 70] <- "60s"
HRMI_a2$CatAge[70 <= HRMI_a2$AGE & HRMI_a2$AGE < 80] <- "70s"
# labeling women and men
HRMI_a2$SEX[HRMI_a2$SEX == 1] <- "Men"
HRMI_a2$SEX[HRMI_a2$SEX == 2] <- "Women"


#data visualization -- distribution of Diabetes by Age and Sex
ggplot(HRMI_a2, aes(fill=SEX, y=DIABETES, 
                    x=factor(CatAge, levels = c("Below 40", "40s","50s", "60s", "70s")))) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual("SEX", values = c("Women" = "pink", "Men" = "light blue")) +
    facet_wrap(~DIABETES) +
    xlab("Age") +
    ylab("Diabete Status") +
    ggtitle("Diabete Status By Age Groups and Sex During Period 1 of the Study") +
  theme( plot.title = element_text(size = 10))
ggsave("Aim2_1.png")

ggplot(HRMI_a2, aes(y = HEARTRTE, x = SEX, color = SEX)) +
  geom_point() +
  facet_wrap(~CatAge) +
  ylab("Count") +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Distribution of Heart Rate in the Study Population By Age Groups and Sex during Period 1") + 
  theme( plot.title = element_text(size = 10))
ggsave("Aim2_2.png")
```
